---
title: "mynah-rhino-interaction"
output:
  pdf_document: default
  html_document: default
date: "2022-11-22"
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
pacman::p_load(dplyr, lme4, MuMIn, ggplot2, coefplot)

before = read.csv('rhino-mynah-before-data.csv')
before$status = as.factor("before")
before.cleaned = before %>%
  filter(!session.no %in% c("1", "4"))

after = read.csv('rhino-mynah-after-data.csv')
after.cleaned = after %>%
  filter(!session.no %in% c("13", "27", "49"))


after.cleaned$status = as.factor('after')
data = rbind(before.cleaned,after.cleaned)

data = data %>%
  rowwise() %>%
  mutate(feeding.birds = sum(feed.on, feed.not.on),
         Time = recode(Time, "2:00 PM" = "14:00", "5:00 PM" = "17:00",
                       "2:15 PM" = "14:15", "2:40 PM" = "14:40", "2:45 PM" = "14:45"),
         Date = gsub('[(Mon)(Tue)(Wed)(Thu)(Fri)]', '', Date),
         datetime = as.POSIXct(paste(Date, Time, sep = " "), format = "%d/%m/%Y %H:%M"),
         hour = lubridate::hour(datetime),
         session.type = recode(hour, "14" = "noon", "17" = "evening"))
# removing Am/Pm from Time
# remove (datename) from data$Date
# Using a datetime instead
# noon vs evening; bird biology

data.cleaned = data %>%
  filter(food.presence == "Yes") %>%
  select(session.no, Weather, rhino.no, food.presence, food.presence.index,
         Interval, feed.on, not.feed.on, feed.not.on, not.feed.not.on,
         mynah.no, status, feeding.birds, datetime, session.type) 

# note: 48 sessions before; 50 sessions after.
```

```{r data manipulation, include=FALSE}
data.analysis = data.cleaned %>%
  group_by(status, session.no) %>%
  dplyr::summarise(total.feed = sum(feeding.birds),
                   total.mynah = sum(mynah.no),
                   prop.feed = total.feed/total.mynah,
         prop.not.feed = 1 - prop.feed,
         max.feed = max(feeding.birds),
         max.birds= max(mynah.no),
         mean.feed = mean(feeding.birds),
         mean.birds = mean(mynah.no),
         weather = first(Weather),
         rhino.no = max(rhino.no),
         food.presence = first(food.presence),
         food.presence.index = max(food.presence.index),
         session.type = first(session.type),
         datetime = first(datetime)) 

# predictors to consider includes: session type (noon vs evening), rhino.no (proxy as food),
# weather (3 levels; cloudy rainy and sunny)

data.analysis$prop.feed[is.na(data.analysis$prop.feed)] = 0
data.analysis$prop.not.feed[is.na(data.analysis$prop.not.feed)] = 0
# because bird counts were done in a minute interval within 20 mins
# it is difficult to justify total bird counts / total feeding birds
# as double-counting (within 20 mins) is quite likely

data.analysis |>
  summarise(cor(max.feed, max.birds),
            cor(max.feed, mean.feed),
            cor(max.birds, mean.birds),
            cor(mean.feed, mean.birds))
# Max no. of feeding birds highly correlated with max no. of birds, even between status
```

```{r analysis via lmer, include = FALSE}
# linear regression as we're using max of no. birds
m1 = lm(max.birds ~ 1,
           data = data.analysis)

m2 = lm(max.birds ~ weather,
           data = data.analysis)

m3 = lm(max.birds ~ rhino.no,
           data = data.analysis)

m4 = lm(max.birds ~ status,
        data = data.analysis)

m5 = lm(max.birds ~ session.type,
        data = data.analysis)

m6 = lm(max.birds ~ status + rhino.no,
        data = data.analysis)

m7 = lm(max.birds ~ status + session.type,
        data = data.analysis)

m8 = lm(max.birds ~ status + weather + session.type,
        data = data.analysis)

m9 = lm(max.birds ~ status + weather + session.type + rhino.no,
        data = data.analysis)

model.sel(m1, m2, m3, m4, m5, m6, m7, m8, m9)
```


```{r summary of m9}
summary(m9)
```

```{r coefficient plots}
coefplot(m9)
```

m9 intercept refers to rhino numbers = 2, during the evening, cloudy weather and before sprinkler installation.

Therefore, m9 suggests that mynah numbers are lowest after the installation of the sprinklers, during the evening, when 2 rhinos are on exhibit and during the cloudy and/or rainy weather (not significant between these two levels). 

So, mynah numbers increases with increasing rhinos (coefficient = 1.8203) on display. That is, every increase of one rhino leads to a increase in maximum number of mynahs by 1.82 (~ to 2).

In other words, sampling in the afternoon also tends to result in a larger number of mynahs present; this has relevance to bird biology

```{r result visualisation}
data.analysis %>%
  group_by(status) %>%
  dplyr::summarise(mean.total = mean(total.mynah),
                   mean.feed = mean(total.feed),
                   max.birds = max(max.feed),
                   max.total= max(max.birds))

# mynah between status
data.analysis %>%
  ggplot(aes(x = status, y = max.birds)) +
  geom_bar(stat = 'identity') + 
  facet_grid(session.type~weather)

# mynah count over time, trend, between status
data.analysis %>%
  ggplot() +
  geom_point(aes(x = session.no, y = max.birds, colour = status))+
  geom_line(aes(x = session.no, y = max.birds, colour = status)) +
  facet_grid(.~status)

```
